---
title: "Pump it up - Predicting faulty pumps"
author: "Arun Marria"
date: "19/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The project uses data from [Taarifa](http://taarifa.org/) and [the Tanzanian Ministry of Water](https://www.drivendata.org/competitions/7/pump-it-up-data-mining-the-water-table/), to predict water which pumps are functional, which need some repairs, and which don't work at all.  I came across this problem and dataset from one of the machine learning competition hosted by [datadriven.org](https://www.drivendata.org/competitions/7/pump-it-up-data-mining-the-water-table/).

## Table of Contents
- [Loading data and required libraries](#Loading data and required packages)


### Loading data and required packages

Loading the libraries which I'll be using through out the project.

```{r, warning=FALSE, message=FALSE}
#loading libraries 
library(dplyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(knitr)
library(repr)

```

Reading data from `csv` files. 

```{r, message=FALSE, results='markup'}

test_df <- read_csv("../data/testFeatures.csv") # test data for ranking in competition
pumps_df <- read_csv("../data/trainFeatures.csv")
temp_df <- read_csv("../data/trainLabels.csv") 
pumps_df <- merge(pumps_df, temp_df, by = "id" ) #actual data for train

kable(pumps_df[1:5,])


```



Inspecting the dataframe. 

```{r}
str(pumps_df)
```

### Visualizing and exploring dataset

After loading dataset, it is time to explore the dataset columns check column types and values and see if they are required for analysis or if they can simply be avoided. 


Before diving into columns, I will make some functions to generate plots and reuse them again and again in my analysis to exlore different variables. 

```{r}
func_boxplot <- function(df, cat_column, numeric_column){
  
  p <- ggplot(df, aes(x ={{cat_column}}, y = {{numeric_column}})) +
    geom_boxplot(fill = "slateblue", alpha =0.2)
  p
  }
```

Checking numeric column `amount_tsh` 

